### Below you will find the code associated with "Marine Protected Areas improve age structure of tropical Pacific tunas" by Jaskiel et al.

### All code and data can be found on GitHub (https://github.com/jaskielj/MPA-Effects-on-Tropical-Tuna-Size-Structure) and Zenodo ()

### Set Directory, Install/library packages, & bring in data
```{r}
setwd("/Users/Jacob/Documents/MPA Effects on Tuna Size Structure")
getwd()

library(tidyverse)
library(ggridges)

length <- read_rds("wcpfc_length_data.rds")
catch_purse_seine <- read_rds("allrfmo_month_1deg_purseseine.rds")
catch_longline <- read_rds("allrfmo_month_5deg_longline_flag.rds")
```

## Exploring Tuna Length in and around MPAs

### Here we'll investigate whether the size distributions of bigeye, yellowfin, and skipjack tunas change after an area is protected. I'll start with a 5 degree buffer around MPAs first, moving onto de-facto MPAs (dfMPAs) and 10 degree buffer zones later on
```{r}
### Starting with PIPA (+5 degree buffer zone)

### PIPA Longline bigeye pre protections
PIPA_LL_bet_length_BP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2014), ### Keep records between beginning of time series (1953) and 2014 (year before protections)
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### PIPA longline bigeye during protection
PIPA_LL_bet_length_DP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2015, 2022), ### Keep records between 2015 and 2022
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_bet_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### PIPA longline bigeye after protections lifted
PIPA_LL_bet_length_AP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2023, 2025), ### Keep records between 2015 and 2022
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting, add in lines for protected status later
PIPA_LL_bet_length <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2025), ### Keep all records
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PIPA_5degbuff_LL_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PIPA_LL_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()

### Trying a different plot, and not quarterly
PIPA_LL_bet_length <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2025), ### Keep all records
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  select(year, species, length_cm)
#group_by(year, species) |>
  #summarise(median_length_cm = median(length_cm, na.rm = TRUE)) |>

ggplot(data = PIPA_LL_bet_length,
       mapping = aes(x = year,
                     y = median_length_cm))
```

### PIPA Bigeye from purse-seine
```{r}
### PIPA (+5 degree buffer) purse-seine bigeye pre protections
PIPA_PS_bet_length_BP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2014), ### Keep records between 1953 and 2014
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### PIPA purse-seine bigeye during protections
PIPA_PS_bet_length_DP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2015, 2022), ### Keep records between 2015 and 2022
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_bet_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### PIPA purse-seine bigeye after protections
PIPA_PS_bet_length_AP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2023, 2025), ### Keep records between 2023 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PIPA_PS_bet_length <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PIPA_5degbuff_PS_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PIPA_PS_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### Yellowfin longline in PIPA (+5 degree buffer)
```{r}
### Yellowfin longline before protections
PIPA_LL_yft_length_BP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2014), ### Keep records between 1953 and 2014
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Yellowfin longline during protections
PIPA_LL_yft_length_DP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2015, 2022), ### Keep records between 2015 and 2022
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_yft_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Yellowfin longline after protections
PIPA_LL_yft_length_AP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2023, 2025), ### Keep records between 2023 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PIPA_LL_yft_length <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PIPA_5degbuff_LL_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PIPA_LL_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### Yellowfin purse-seine in PIPA (+5 degree buffer)
```{r}
### Yellowfin purse-seine before protections
PIPA_PS_yft_length_BP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2014), ### Keep records between 1953 and 2014
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Yellowfin purse-seine during protections
PIPA_PS_yft_length_DP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2015, 2022), ### Keep records between 2015 and 2022
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_yft_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Yellowfin purse-seine after protections
PIPA_PS_yft_length_AP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2023, 2025), ### Keep records between 2023 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PIPA_PS_yft_length <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PIPA_5degbuff_PS_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PIPA_PS_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### Skipjack longline (not major component but still looking) in PIPA (+5 degree buffer)
```{r}
### Skipjack longline before protections
PIPA_LL_skj_length_BP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2014), ### Keep records between 1953 and 2014
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Skipjack longline during protections
PIPA_LL_skj_length_DP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2015, 2022), ### Keep records between 2015 and 2022
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_skj_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Skipjack longline after protections
PIPA_LL_skj_length_AP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2023, 2025), ### Keep records between 2023 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PIPA_LL_skj_length <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PIPA_5degbuff_LL_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PIPA_LL_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### Skipjack purse-seine in PIPA (+5 degree buffer)
```{r}
### Skipjack purse-seine before protections
PIPA_PS_skj_length_BP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2014), ### Keep records between 1953 and 2014
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Skipjack purse-seine during protections
PIPA_PS_skj_length_DP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2015, 2022), ### Keep records between 2015 and 2022
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_skj_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Skipjack purse-seine after protections
PIPA_PS_skj_length_AP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2023, 2025), ### Keep records between 2023 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PIPA_PS_skj_length <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PIPA_5degbuff_PS_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PIPA_PS_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```


## Palau National Marine Sanctuary (+5 degree buffer)

### PNMS Bigeye
```{r}
### Longline pre-protection
PNMS_LL_bet_length_BP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_LL_bet_length_AP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PNMS_LL_bet_length <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PNMS_5degbuff_LL_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PNMS_LL_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### PNMS purse-seine Bigeye
```{r}
### Before protection
PNMS_PS_bet_length_BP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_PS_bet_length_AP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PNMS_PS_bet_length <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PNMS_5degbuff_PS_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PNMS_PS_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### PNMS Yellowfin longline
```{r}
### Before protection
PNMS_LL_yft_length_BP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_LL_yft_length_AP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PNMS_LL_yft_length <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PNMS_5degbuff_LL_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PNMS_LL_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### PNMS purse-seine yellowfin
```{r}
### Before protection
PNMS_PS_yft_length_BP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_PS_yft_length_AP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PNMS_PS_yft_length <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PNMS_5degbuff_PS_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PNMS_PS_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### PNMS longline skipjack
```{r}
### Before protection
PNMS_LL_skj_length_BP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_LL_skj_length_AP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PNMS_LL_skj_length <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PNMS_5degbuff_LL_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PNMS_LL_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### PNMS purse-seine skipjack
```{r}
### Before protection
PNMS_PS_skj_length_BP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_PS_skj_length_AP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PNMS_PS_skj_length <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PNMS_5degbuff_PS_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PNMS_PS_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

## Galapagos MPA
###Galapagos purse seine bigeye with 5 degree buffer (no LL data within these spatial constraints)
```{r}
### Pre-protections
GMR_PS_bet_length_BP <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1953, 1997), ### Keep records between 1953 and 1997
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = GMR_PS_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After Protections
GMR_PS_bet_length_AP <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1998, 2025), ### Keep records between 1998 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = GMR_PS_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
GMR_PS_bet_length <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("GMR_5degbuff_PS_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = GMR_PS_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### Galapagos (+5 degree buffer) yellowfin purse seine (no longline data)
```{r}
### Before protections
GMR_PS_yft_length_BP <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1953, 1997), ### Keep records between 1953 and 1997
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = GMR_PS_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protections
GMR_PS_yft_length_AP <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1998, 2025), ### Keep records between 1998 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = GMR_PS_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
GMR_PS_yft_length <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("GMR_5degbuff_PS_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = GMR_PS_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### GMR (+5 degree buffer) skipjack purse seine (no longline data)
```{r}
### Before protections
GMR_PS_skj_length_BP <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1953, 1997), ### Keep records between 1953 and 1997
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = GMR_PS_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After Protections
GMR_PS_skj_length_AP <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1998, 2025), ### Keep records between 1998 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = GMR_PS_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
GMR_PS_skj_length <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("GMR_5degbuff_PS_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = GMR_PS_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

## Papahānaumokuākea 
### Papahānaumokuākea (+5 degree buffer) longline - bigeye (note: can't cross international dateline with this version so set min lon to -180)
```{r}
### Before protections
PMNM_LL_bet_length_BP <- length |>
  filter(between(lon, -180, -156), ### Filter around PMNM + 5 degree buffer
         between(lat, 14.25, 37),
         between(year, 1953, 2015), ### Keep records between 1953 and 2015
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PMNM_LL_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protections
PMNM_LL_bet_length_AP <- length |>
  filter(between(lon, -180, -156), ### Filter around PMNM + 5 degree buffer
         between(lat, 14.25, 37),
         between(year, 2016, 2025), ### Keep records between 2016 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PMNM_LL_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PMNM_LL_bet_length <- length |>
  filter(between(lon, -180, -156), ### Filter around PMNM + 5 degree buffer
         between(lat, 14.25, 37),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PMNM_5degbuff_LL_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PMNM_LL_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### PMNM (+5 degree buffer) longline - yellowfin (no data for purse seine bigeye except 1998)
```{r}
### Before protections
PMNM_LL_yft_length_BP <- length |>
  filter(between(lon, -180, -156), ### Filter around PMNM + 5 degree buffer
         between(lat, 14.25, 37),
         between(year, 1953, 2015), ### Keep records between 1953 and 2015
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PMNM_LL_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protections
PMNM_LL_yft_length_AP <- length |>
  filter(between(lon, -180, -156), ### Filter around PMNM + 5 degree buffer
         between(lat, 14.25, 37),
         between(year, 2016, 2025), ### Keep records between 2016 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PMNM_LL_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PMNM_LL_yft_length <- length |>
  filter(between(lon, -180, -156), ### Filter around PMNM + 5 degree buffer
         between(lat, 14.25, 37),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PMNM_5degbuff_LL_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PMNM_LL_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```
### Note: no data for skipjack for longline or purse seine for this region oddly...
-22.5, -17.25
-170, -167


## Niue - Moana Mahu Marine Protected Area
### MMMPA (+5 degree buffer) longline - bigeye
```{r}
### Before protections
MMMPA_LL_bet_length_BP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_LL_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protections
MMMPA_LL_bet_length_AP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_LL_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
MMMPA_LL_bet_length <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("MMMPA_5degbuff_LL_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = MMMPA_LL_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### MMMPA purse seine - bigeye
```{r}
### Before protections
MMMPA_PS_bet_length_BP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_PS_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protections
MMMPA_PS_bet_length_AP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_PS_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
MMMPA_PS_bet_length <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("MMMPA_5degbuff_PS_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = MMMPA_PS_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### MMMPA (+5 degree buffer) longline - yellowfin
```{r}
### Before protections
MMMPA_LL_yft_length_BP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_LL_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protections
MMMPA_LL_yft_length_AP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_LL_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
MMMPA_LL_yft_length <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("MMMPA_5degbuff_LL_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = MMMPA_LL_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```


### MMMPA (+5 degree buffer) purse seine - yellowfin
```{r}
### Before protections
MMMPA_PS_yft_length_BP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_PS_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protections
MMMPA_PS_yft_length_AP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_PS_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
MMMPA_PS_yft_length <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("MMMPA_5degbuff_PS_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = MMMPA_PS_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### MMMPA (+5 degree buffer) longline - skipjack
```{r}
### Before protections
MMMPA_LL_skj_length_BP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_LL_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protections
MMMPA_LL_skj_length_AP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_LL_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
MMMPA_LL_skj_length <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("MMMPA_5degbuff_LL_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = MMMPA_LL_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### MMMPA (+5 degree buffer) purse-seine - skipjack
```{r}
### Before protections
MMMPA_PS_skj_length_BP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_PS_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protections
MMMPA_PS_skj_length_AP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_PS_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
MMMPA_PS_skj_length <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("MMMPA_5degbuff_PS_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = MMMPA_PS_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

## High Seas Pocket (no buffer)
### High Seas Pocket (no buffer) longline - bigeye
```{r}
### Whole time series
HSP_LL_bet_length <- length |>
  filter(between(lon, 136, 152), ### Filter around HSP
         between(lat, 1, 5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("HSP_LL_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = HSP_LL_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### High Seas Pocket (no buffer) purse seine - bigeye
```{r}
### Whole time series
HSP_PS_bet_length <- length |>
  filter(between(lon, 136, 152), ### Filter around HSP
         between(lat, 1, 5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("HSP_PS_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = HSP_PS_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### High Seas Pocket (no buffer) longline - yellowfin
```{r}
### Whole time series
HSP_LL_yft_length <- length |>
  filter(between(lon, 136, 152), ### Filter around HSP
         between(lat, 1, 5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("HSP_LL_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = HSP_LL_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### High Seas Pocket (no buffer) purse seine - yellowfin
```{r}
### Whole time series
HSP_PS_yft_length <- length |>
  filter(between(lon, 136, 152), ### Filter around HSP
         between(lat, 1, 5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("HSP_PS_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = HSP_PS_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### High Seas Pocket (no buffer) longline - skipjack
```{r}
### Whole time series
HSP_LL_skj_length <- length |>
  filter(between(lon, 136, 152), ### Filter around HSP
         between(lat, 1, 5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("HSP_LL_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = HSP_LL_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### High Seas Pocket (no buffer) purse seine - skipjack
```{r}
### Whole time series
HSP_PS_skj_length <- length |>
  filter(between(lon, 136, 152), ### Filter around HSP
         between(lat, 1, 5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("HSP_PS_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = HSP_PS_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### Adding some scripts from Lynham and Villasenor-Derbez (2024) to use a spatial grid rather than what I've been using, which has been coarse latitude and longitude-based analyses that didn't allow for precise calculation of changes in tuna length near (<100 nm) and far (100-200 nm) from each MPA
```{r}
### downloaded 00_set_up.R from https://github.com/jcvdav/LSMPA_spillover/blob/main
## SET UP ######################################################################
#install.packages("pacman")

# Load packages ----------------------------------------------------------------
pacman::p_load(
  magrittr,
  here,
  terra,
  sf,
  units,
  tidyverse
)

# Source custom functions ------------------------------------------------------
source("/Users/Jacob/Documents/MPA Effects on Tuna Size Structure/00_set_up.R")

# Define functions -------------------------------------------------------------
# This function assigns a "near" or "far" category to each point, depending on
# the value of x and the distance to an MPA
assign_near_far <- function(dist, x){
  case_when((dist >= 0 & dist <= x) ~ "near",
            (dist > x & dist <= (2 * x)) ~ "far",
            T ~ NA_character_)
}


###FLAG: stopped here 10/7

# Load data --------------------------------------------------------------------
data <- readRDS(file = here("data", "processed", "rfmo_all_annual_gear_flag.rds"))
lmpas <- st_read(dsn = here("data", "processed", "clean_lmpas.gpkg"))


## PROCESSING ##################################################################

# Find cells covered by MPAs ---------------------------------------------

# First, build some rasters
cells_1x1_within <-
  rast(nrows = 180, ncols = 360,
       xmin = -180, xmax = 180,
       ymin = -90, ymax = 90,
       nlyrs = 1, crs = "EPSG: 4326",
       resolution = 1, val = 1) %>%
  # Identify which cells are within an MPA polygon
  terra::extract(lmpas,
          xy = T,
          exact = T)  %>% # The `exact = T` argument calculates the percent cover
  mutate(grid = "1x1") %>%
  select(grid, lon = x, lat = y, fraction)

dim(cells_1x1_within) # There are 914 (781 without Johsnton) cells on the 1X1 grid that fall
# within an MPA. Note that not all these cells may have data reported by
# an RFMO and that not all are FULLY within an MPA... we'll get there

cells_5x5_within <-
  rast(nrows = 180, ncols = 360,
       xmin = -180, xmax = 180,
       ymin = -90, ymax = 90,
       nlyrs = 1, crs = "EPSG: 4326",
       resolution = 5, val = 1) %>%
  # Identify which cells are within an MPA polygon
  terra::extract(lmpas,
          xy = T,
          exact = T) %>%
  mutate(grid = "5x5") %>%
  select(grid, lon = x, lat = y, fraction)

dim(cells_5x5_within) # There are 97 (81 without Johnston) cells within an MPA using the 5x5 grid


cells_within <- bind_rows(cells_1x1_within,
                          cells_5x5_within) %>%
  filter(fraction >= 1) %>%
  select(-fraction)

# Build spatial grid -----------------------------------------------------------
# There are a few steps here:
# - First, build a grid of unique lat-lon combinations where RFMOs actually report
# some data
# - Then, remove cells that we've identified as being FULLY covered by an MPA
# - We then calculate the distance of each point to the nearest MP

# Create a single feature to keep things tidy
lmpas_sfc <- lmpas %>%
  st_union()

# Make grid with distances and donuts ------------------------------------------
rfmo_spat_grid <- data %>%
  # Build distinct lat, lon pairs by grid type
  select(grid, lat, lon) %>%
  distinct() %>%
  # Remove cells identified as fully within an MPA
  anti_join(cells_within, by = c("grid", "lat", "lon")) %>% # This removes 110 grid cells that fall within an MPA and that have data
  # Convert into a spatial object
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  # Calculate distance to nearest MPA
  mutate(dist = st_distance(., lmpas_sfc),
         dist = set_units(x = dist, value = "nautical_miles"), # From now on, distances are in nautical mules
         dist = as.numeric(dist)) %>%
  # We'll keep only grid cells within 600 nautical miles of an MPA
  filter(dist <= 600) %>%
  # Build a series of dummies
  mutate(treatment_100 = assign_near_far(dist = dist, x = 100),
         treatment_200 = assign_near_far(dist = dist, x = 200),
         treatment_300 = assign_near_far(dist = dist, x = 300),
         near_100 = case_when(treatment_100 == "near" ~ 1,
                              treatment_100 == "far" ~ 0,
                              T ~ NA),
         near_200 = case_when(treatment_200 == "near" ~ 1,
                              treatment_200 == "far" ~ 0,
                              T ~ NA),
         near_300 = case_when(treatment_300 == "near" ~ 1,
                              treatment_300 == "far" ~ 0,
                              T ~ NA)) %>%
  # We now add information on which one is the nearest MPA
  st_join(lmpas, st_nearest_feature)


# Now convert it into a data frame
rfmo_grid <- rfmo_spat_grid %>%
  bind_cols(st_coordinates(.)) %>%
  st_drop_geometry() %>%
  rename(lon = X, lat = Y) %>%
  group_by(lat, lon) %>%
  mutate(id = cur_group_id()) %>%
  ungroup() %>%
  select(id, grid, lat, lon, wdpaid, name, dist, contains("treatment"), contains("near_"), ocean, year_enforced, month_enforced)

## EXPORT ######################################################################
saveRDS(object = rfmo_grid,
        file = here("data", "processed", "distance_grid.rds"))
```

