### Below you will find the code associated with "Marine Protected Areas improve age structure of tropical Pacific tunas" by Jaskiel et al.

### All code and data can be found on GitHub (https://github.com/jaskielj/MPA-Effects-on-Tropical-Tuna-Size-Structure) and Zenodo ()

### Set Directory, Install/library packages, & bring in data
```{r}
setwd("/Users/Jacob/Documents/MPA Effects on Tuna Size Structure")
getwd()

library(tidyverse)
library(ggridges)

length <- read_rds("wcpfc_length_data.rds")
catch_purse_seine <- read_rds("allrfmo_month_1deg_purseseine.rds")
catch_longline <- read_rds("allrfmo_month_5deg_longline_flag.rds")
```

## Exploring Tuna Length in and around MPAs

### Here we'll investigate whether the size distributions of bigeye, yellowfin, and skipjack tunas change after an area is protected. I'll start with a 5 degree buffer around MPAs first, moving onto de-facto MPAs (dfMPAs) and 10 degree buffer zones later on
```{r}
### Starting with PIPA (+5 degree buffer zone)

### PIPA Longline bigeye pre protections
PIPA_LL_bet_length_BP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2014), ### Keep records between beginning of time series (1953) and 2014 (year before protections)
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### PIPA longline bigeye during protection
PIPA_LL_bet_length_DP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2015, 2022), ### Keep records between 2015 and 2022
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_bet_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### PIPA longline bigeye after protections lifted
PIPA_LL_bet_length_AP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2023, 2025), ### Keep records between 2015 and 2022
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting, add in lines for protected status later
PIPA_LL_bet_length <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2025), ### Keep all records
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PIPA_5degbuff_LL_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PIPA_LL_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### PIPA Bigeye from purse-seine
```{r}
### PIPA (+5 degree buffer) purse-seine bigeye pre protections
PIPA_PS_bet_length_BP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2014), ### Keep records between 1953 and 2014
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### PIPA purse-seine bigeye during protections
PIPA_PS_bet_length_DP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2015, 2022), ### Keep records between 2015 and 2022
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_bet_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### PIPA purse-seine bigeye after protections
PIPA_PS_bet_length_AP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2023, 2025), ### Keep records between 2023 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PIPA_PS_bet_length <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PIPA_5degbuff_PS_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PIPA_PS_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### Yellowfin longline in PIPA (+5 degree buffer)
```{r}
### Yellowfin longline before protections
PIPA_LL_yft_length_BP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2014), ### Keep records between 1953 and 2014
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Yellowfin longline during protections
PIPA_LL_yft_length_DP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2015, 2022), ### Keep records between 2015 and 2022
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_yft_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Yellowfin longline after protections
PIPA_LL_yft_length_AP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2023, 2025), ### Keep records between 2023 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PIPA_LL_yft_length <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PIPA_5degbuff_LL_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PIPA_LL_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### Yellowfin purse-seine in PIPA (+5 degree buffer)
```{r}
### Yellowfin purse-seine before protections
PIPA_PS_yft_length_BP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2014), ### Keep records between 1953 and 2014
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Yellowfin purse-seine during protections
PIPA_PS_yft_length_DP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2015, 2022), ### Keep records between 2015 and 2022
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_yft_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Yellowfin purse-seine after protections
PIPA_PS_yft_length_AP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2023, 2025), ### Keep records between 2023 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PIPA_PS_yft_length <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PIPA_5degbuff_PS_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PIPA_PS_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### Skipjack longline (not major component but still looking) in PIPA (+5 degree buffer)
```{r}
### Skipjack longline before protections
PIPA_LL_skj_length_BP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2014), ### Keep records between 1953 and 2014
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Skipjack longline during protections
PIPA_LL_skj_length_DP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2015, 2022), ### Keep records between 2015 and 2022
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_skj_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Skipjack longline after protections
PIPA_LL_skj_length_AP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2023, 2025), ### Keep records between 2023 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PIPA_LL_skj_length <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PIPA_5degbuff_LL_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PIPA_LL_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### Skipjack purse-seine in PIPA (+5 degree buffer)
```{r}
### Skipjack purse-seine before protections
PIPA_PS_skj_length_BP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2014), ### Keep records between 1953 and 2014
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Skipjack purse-seine during protections
PIPA_PS_skj_length_DP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2015, 2022), ### Keep records between 2015 and 2022
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_skj_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Skipjack purse-seine after protections
PIPA_PS_skj_length_AP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2023, 2025), ### Keep records between 2023 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PIPA_PS_skj_length <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PIPA_5degbuff_PS_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PIPA_PS_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```


## Palau National Marine Sanctuary (+5 degree buffer)

### PNMS Bigeye
```{r}
### Longline pre-protection
PNMS_LL_bet_length_BP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_LL_bet_length_AP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PNMS_LL_bet_length <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PNMS_5degbuff_LL_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PNMS_LL_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### PNMS purse-seine Bigeye
```{r}
### Before protection
PNMS_PS_bet_length_BP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_PS_bet_length_AP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PNMS_PS_bet_length <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PNMS_5degbuff_PS_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PNMS_PS_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### PNMS Yellowfin longline
```{r}
### Before protection
PNMS_LL_yft_length_BP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_LL_yft_length_AP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PNMS_LL_yft_length <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PNMS_5degbuff_LL_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PNMS_LL_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### PNMS purse-seine yellowfin
```{r}
### Before protection
PNMS_PS_yft_length_BP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_PS_yft_length_AP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PNMS_PS_yft_length <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PNMS_5degbuff_PS_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PNMS_PS_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### PNMS longline skipjack
```{r}
### Before protection
PNMS_LL_skj_length_BP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_LL_skj_length_AP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PNMS_LL_skj_length <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PNMS_5degbuff_LL_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PNMS_LL_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### PNMS purse-seine skipjack
```{r}
### Before protection
PNMS_PS_skj_length_BP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_PS_skj_length_AP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PNMS_PS_skj_length <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PNMS_5degbuff_PS_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PNMS_PS_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

## Galapagos MPA
###Galapagos purse seine bigeye with 5 degree buffer (no LL data within these spatial constraints)
```{r}
### Pre-protections
GMR_PS_bet_length_BP <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1953, 1997), ### Keep records between 1953 and 1997
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = GMR_PS_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After Protections
GMR_PS_bet_length_AP <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1998, 2025), ### Keep records between 1998 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = GMR_PS_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
GMR_PS_bet_length <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("GMR_5degbuff_PS_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = GMR_PS_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### Galapagos (+5 degree buffer) yellowfin purse seine (no longline data)
```{r}
### Before protections
GMR_PS_yft_length_BP <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1953, 1997), ### Keep records between 1953 and 1997
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = GMR_PS_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protections
GMR_PS_yft_length_AP <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1998, 2025), ### Keep records between 1998 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = GMR_PS_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
GMR_PS_yft_length <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("GMR_5degbuff_PS_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = GMR_PS_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### GMR (+5 degree buffer) skipjack purse seine (no longline data)
```{r}
### Before protections
GMR_PS_skj_length_BP <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1953, 1997), ### Keep records between 1953 and 1997
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = GMR_PS_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After Protections
GMR_PS_skj_length_AP <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1998, 2025), ### Keep records between 1998 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = GMR_PS_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
GMR_PS_skj_length <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("GMR_5degbuff_PS_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = GMR_PS_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

