### Below you will find the code associated with "Marine Protected Areas improve age structure of tropical Pacific tunas" by Jaskiel et al.

### All code and data can be found on GitHub (https://github.com/jaskielj/MPA-Effects-on-Tropical-Tuna-Size-Structure) and Zenodo ()

### Set Directory, Install/library packages, & bring in data
```{r}
setwd("/Users/Jacob/Documents/MPA Effects on Tuna Size Structure")
getwd()

library(tidyverse)
library(ggridges)

length <- read_rds("wcpfc_length_data.rds")
catch_purse_seine <- read_rds("allrfmo_month_1deg_purseseine.rds")
catch_longline <- read_rds("allrfmo_month_5deg_longline_flag.rds")
```

## Exploring Tuna Length in and around MPAs

### Here we'll investigate whether the size distributions of bigeye, yellowfin, and skipjack tunas change after an area is protected. I'll start with a 5 degree buffer around MPAs first, moving onto de-facto MPAs (dfMPAs) and 10 degree buffer zones later on
```{r}
### Starting with PIPA (+5 degree buffer zone)

### PIPA Longline bigeye pre protections
PIPA_LL_bet_length_BP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2014), ### Keep records between beginning of time series (1953) and 2014 (year before protections)
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### PIPA longline bigeye during protection
PIPA_LL_bet_length_DP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2015, 2022), ### Keep records between 2015 and 2022
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_bet_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### PIPA longline bigeye after protections lifted
PIPA_LL_bet_length_AP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2023, 2025), ### Keep records between 2015 and 2022
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting, add in lines for protected status later
PIPA_LL_bet_length <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2025), ### Keep all records
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PIPA_5degbuff_LL_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PIPA_LL_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()

### Trying a different plot, and not quarterly
PIPA_LL_bet_length <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2025), ### Keep all records
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  select(year, species, length_cm)
#group_by(year, species) |>
  #summarise(median_length_cm = median(length_cm, na.rm = TRUE)) |>

ggplot(data = PIPA_LL_bet_length,
       mapping = aes(x = year,
                     y = median_length_cm))
```

### PIPA Bigeye from purse-seine
```{r}
### PIPA (+5 degree buffer) purse-seine bigeye pre protections
PIPA_PS_bet_length_BP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2014), ### Keep records between 1953 and 2014
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### PIPA purse-seine bigeye during protections
PIPA_PS_bet_length_DP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2015, 2022), ### Keep records between 2015 and 2022
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_bet_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### PIPA purse-seine bigeye after protections
PIPA_PS_bet_length_AP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2023, 2025), ### Keep records between 2023 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PIPA_PS_bet_length <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PIPA_5degbuff_PS_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PIPA_PS_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### Yellowfin longline in PIPA (+5 degree buffer)
```{r}
### Yellowfin longline before protections
PIPA_LL_yft_length_BP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2014), ### Keep records between 1953 and 2014
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Yellowfin longline during protections
PIPA_LL_yft_length_DP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2015, 2022), ### Keep records between 2015 and 2022
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_yft_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Yellowfin longline after protections
PIPA_LL_yft_length_AP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2023, 2025), ### Keep records between 2023 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PIPA_LL_yft_length <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PIPA_5degbuff_LL_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PIPA_LL_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### Yellowfin purse-seine in PIPA (+5 degree buffer)
```{r}
### Yellowfin purse-seine before protections
PIPA_PS_yft_length_BP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2014), ### Keep records between 1953 and 2014
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Yellowfin purse-seine during protections
PIPA_PS_yft_length_DP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2015, 2022), ### Keep records between 2015 and 2022
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_yft_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Yellowfin purse-seine after protections
PIPA_PS_yft_length_AP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2023, 2025), ### Keep records between 2023 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PIPA_PS_yft_length <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PIPA_5degbuff_PS_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PIPA_PS_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### Skipjack longline (not major component but still looking) in PIPA (+5 degree buffer)
```{r}
### Skipjack longline before protections
PIPA_LL_skj_length_BP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2014), ### Keep records between 1953 and 2014
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Skipjack longline during protections
PIPA_LL_skj_length_DP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2015, 2022), ### Keep records between 2015 and 2022
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_skj_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Skipjack longline after protections
PIPA_LL_skj_length_AP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2023, 2025), ### Keep records between 2023 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PIPA_LL_skj_length <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PIPA_5degbuff_LL_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PIPA_LL_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### Skipjack purse-seine in PIPA (+5 degree buffer)
```{r}
### Skipjack purse-seine before protections
PIPA_PS_skj_length_BP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2014), ### Keep records between 1953 and 2014
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Skipjack purse-seine during protections
PIPA_PS_skj_length_DP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2015, 2022), ### Keep records between 2015 and 2022
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_skj_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Skipjack purse-seine after protections
PIPA_PS_skj_length_AP <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 2023, 2025), ### Keep records between 2023 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PIPA_PS_skj_length <- length |>
  filter(between(lon, -179.999, -164.706), ### Filter around PIPA + 5 degree buffer
         between(lat, -11.618, 4.983),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PIPA_5degbuff_PS_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PIPA_PS_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```


## Palau National Marine Sanctuary (+5 degree buffer)

### PNMS Bigeye
```{r}
### Longline pre-protection
PNMS_LL_bet_length_BP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_LL_bet_length_AP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PNMS_LL_bet_length <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PNMS_5degbuff_LL_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PNMS_LL_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### PNMS purse-seine Bigeye
```{r}
### Before protection
PNMS_PS_bet_length_BP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_PS_bet_length_AP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PNMS_PS_bet_length <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PNMS_5degbuff_PS_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PNMS_PS_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### PNMS Yellowfin longline
```{r}
### Before protection
PNMS_LL_yft_length_BP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_LL_yft_length_AP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PNMS_LL_yft_length <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PNMS_5degbuff_LL_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PNMS_LL_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### PNMS purse-seine yellowfin
```{r}
### Before protection
PNMS_PS_yft_length_BP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_PS_yft_length_AP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PNMS_PS_yft_length <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PNMS_5degbuff_PS_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PNMS_PS_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### PNMS longline skipjack
```{r}
### Before protection
PNMS_LL_skj_length_BP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_LL_skj_length_AP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PNMS_LL_skj_length <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PNMS_5degbuff_LL_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PNMS_LL_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### PNMS purse-seine skipjack
```{r}
### Before protection
PNMS_PS_skj_length_BP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_PS_skj_length_AP <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PNMS_PS_skj_length <- length |>
  filter(between(lon, 124.5, 142), ### Filter around PNMS w/ 5 degree buffer
         between(lat, -3.5, 16.5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PNMS_5degbuff_PS_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PNMS_PS_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

## Galapagos MPA
###Galapagos purse seine bigeye with 5 degree buffer (no LL data within these spatial constraints)
```{r}
### Pre-protections
GMR_PS_bet_length_BP <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1953, 1997), ### Keep records between 1953 and 1997
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = GMR_PS_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After Protections
GMR_PS_bet_length_AP <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1998, 2025), ### Keep records between 1998 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = GMR_PS_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
GMR_PS_bet_length <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("GMR_5degbuff_PS_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = GMR_PS_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### Galapagos (+5 degree buffer) yellowfin purse seine (no longline data)
```{r}
### Before protections
GMR_PS_yft_length_BP <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1953, 1997), ### Keep records between 1953 and 1997
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = GMR_PS_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protections
GMR_PS_yft_length_AP <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1998, 2025), ### Keep records between 1998 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = GMR_PS_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
GMR_PS_yft_length <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("GMR_5degbuff_PS_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = GMR_PS_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### GMR (+5 degree buffer) skipjack purse seine (no longline data)
```{r}
### Before protections
GMR_PS_skj_length_BP <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1953, 1997), ### Keep records between 1953 and 1997
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = GMR_PS_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After Protections
GMR_PS_skj_length_AP <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1998, 2025), ### Keep records between 1998 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = GMR_PS_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
GMR_PS_skj_length <- length |>
  filter(between(lon, -97.681, -83.4989), ### Filter around GMR + 5 degree buffer
         between(lat, -7.1135, 7.3534),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("GMR_5degbuff_PS_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = GMR_PS_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

## Papahānaumokuākea 
### Papahānaumokuākea (+5 degree buffer) longline - bigeye (note: can't cross international dateline with this version so set min lon to -180)
```{r}
### Before protections
PMNM_LL_bet_length_BP <- length |>
  filter(between(lon, -180, -156), ### Filter around PMNM + 5 degree buffer
         between(lat, 14.25, 37),
         between(year, 1953, 2015), ### Keep records between 1953 and 2015
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PMNM_LL_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protections
PMNM_LL_bet_length_AP <- length |>
  filter(between(lon, -180, -156), ### Filter around PMNM + 5 degree buffer
         between(lat, 14.25, 37),
         between(year, 2016, 2025), ### Keep records between 2016 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PMNM_LL_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PMNM_LL_bet_length <- length |>
  filter(between(lon, -180, -156), ### Filter around PMNM + 5 degree buffer
         between(lat, 14.25, 37),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PMNM_5degbuff_LL_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PMNM_LL_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### PMNM (+5 degree buffer) longline - yellowfin (no data for purse seine bigeye except 1998)
```{r}
### Before protections
PMNM_LL_yft_length_BP <- length |>
  filter(between(lon, -180, -156), ### Filter around PMNM + 5 degree buffer
         between(lat, 14.25, 37),
         between(year, 1953, 2015), ### Keep records between 1953 and 2015
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PMNM_LL_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protections
PMNM_LL_yft_length_AP <- length |>
  filter(between(lon, -180, -156), ### Filter around PMNM + 5 degree buffer
         between(lat, 14.25, 37),
         between(year, 2016, 2025), ### Keep records between 2016 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PMNM_LL_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
PMNM_LL_yft_length <- length |>
  filter(between(lon, -180, -156), ### Filter around PMNM + 5 degree buffer
         between(lat, 14.25, 37),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("PMNM_5degbuff_LL_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = PMNM_LL_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```
### Note: no data for skipjack for longline or purse seine for this region oddly...
-22.5, -17.25
-170, -167


## Niue - Moana Mahu Marine Protected Area
### MMMPA (+5 degree buffer) longline - bigeye
```{r}
### Before protections
MMMPA_LL_bet_length_BP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_LL_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protections
MMMPA_LL_bet_length_AP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_LL_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
MMMPA_LL_bet_length <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("MMMPA_5degbuff_LL_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = MMMPA_LL_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### MMMPA purse seine - bigeye
```{r}
### Before protections
MMMPA_PS_bet_length_BP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_PS_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protections
MMMPA_PS_bet_length_AP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_PS_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
MMMPA_PS_bet_length <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("MMMPA_5degbuff_PS_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = MMMPA_PS_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### MMMPA (+5 degree buffer) longline - yellowfin
```{r}
### Before protections
MMMPA_LL_yft_length_BP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_LL_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protections
MMMPA_LL_yft_length_AP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_LL_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
MMMPA_LL_yft_length <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("MMMPA_5degbuff_LL_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = MMMPA_LL_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```


### MMMPA (+5 degree buffer) purse seine - yellowfin
```{r}
### Before protections
MMMPA_PS_yft_length_BP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_PS_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protections
MMMPA_PS_yft_length_AP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_PS_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
MMMPA_PS_yft_length <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("MMMPA_5degbuff_PS_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = MMMPA_PS_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### MMMPA (+5 degree buffer) longline - skipjack
```{r}
### Before protections
MMMPA_LL_skj_length_BP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_LL_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protections
MMMPA_LL_skj_length_AP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_LL_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
MMMPA_LL_skj_length <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("MMMPA_5degbuff_LL_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = MMMPA_LL_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### MMMPA (+5 degree buffer) purse-seine - skipjack
```{r}
### Before protections
MMMPA_PS_skj_length_BP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2019), ### Keep records between 1953 and 2019
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_PS_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protections
MMMPA_PS_skj_length_AP <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 2020, 2025), ### Keep records between 2020 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = MMMPA_PS_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Whole time series for plotting
MMMPA_PS_skj_length <- length |>
  filter(between(lon, -175, -162), ### Filter around MMMPA + 5 degree buffer
         between(lat, -27.5, -12.25),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("MMMPA_5degbuff_PS_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = MMMPA_PS_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

## High Seas Pocket (no buffer)
### High Seas Pocket (no buffer) longline - bigeye
```{r}
### Whole time series
HSP_LL_bet_length <- length |>
  filter(between(lon, 136, 152), ### Filter around HSP
         between(lat, 1, 5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("HSP_LL_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = HSP_LL_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### High Seas Pocket (no buffer) purse seine - bigeye
```{r}
### Whole time series
HSP_PS_bet_length <- length |>
  filter(between(lon, 136, 152), ### Filter around HSP
         between(lat, 1, 5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("HSP_PS_bet_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = HSP_PS_bet_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### High Seas Pocket (no buffer) longline - yellowfin
```{r}
### Whole time series
HSP_LL_yft_length <- length |>
  filter(between(lon, 136, 152), ### Filter around HSP
         between(lat, 1, 5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("HSP_LL_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = HSP_LL_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### High Seas Pocket (no buffer) purse seine - yellowfin
```{r}
### Whole time series
HSP_PS_yft_length <- length |>
  filter(between(lon, 136, 152), ### Filter around HSP
         between(lat, 1, 5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "YFT", ### Keep yellowfin tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("HSP_PS_yft_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = HSP_PS_yft_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### High Seas Pocket (no buffer) longline - skipjack
```{r}
### Whole time series
HSP_LL_skj_length <- length |>
  filter(between(lon, 136, 152), ### Filter around HSP
         between(lat, 1, 5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("HSP_LL_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = HSP_LL_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```

### High Seas Pocket (no buffer) purse seine - skipjack
```{r}
### Whole time series
HSP_PS_skj_length <- length |>
  filter(between(lon, 136, 152), ### Filter around HSP
         between(lat, 1, 5),
         between(year, 1953, 2025), ### Keep records between 1953 and 2025
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on purse seine
  mutate(yq = paste(year, quarter)) |>
  select(yq, species, length_cm)

### Plot
#png("HSP_PS_skj_length.png", units = "in", width = 8, height = 5, res = 1200)
ggplot(data = HSP_PS_skj_length,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()
```





############################################################################################################################
############################################################################################################################
############################################################################################################################
################################################# PART TWO - LOCKING IN ####################################################
############################################################################################################################
############################################################################################################################
############################################################################################################################

### Adding some scripts from Lynham and Villasenor-Derbez (2024) to use a spatial grid rather than what I've been using, which has been coarse latitude and longitude-based analyses that didn't allow for precise calculation of changes in tuna length near (<100 nm) and far (100-200 nm) from each MPA
```{r}
## SET UP ######################################################################
#install.packages("pacman")
library(pacman)

# Load packages ----------------------------------------------------------------
pacman::p_load(
  magrittr,
  here,
  terra,
  sf,
  units,
  tidyverse
)

# Source custom functions ------------------------------------------------------
### downloaded 00_set_up.R from https://github.com/jcvdav/LSMPA_spillover/blob/main
### This document has necessary custom functions and color pallettes, amongst other things
source("00_set_up.R") 

# Define functions -------------------------------------------------------------
# This function assigns a "near" or "far" category to each point, depending on
# the value of x and the distance to an MPA
assign_near_far <- function(dist, x){
  case_when((dist >= 0 & dist <= x) ~ "near",
            (dist > x & dist <= (2 * x)) ~ "far",
            T ~ NA_character_)
}

# Load data --------------------------------------------------------------------
length <- read_rds("wcpfc_length_data.rds")
catch_purse_seine <- read_rds("allrfmo_month_1deg_purseseine.rds")
catch_longline <- read_rds("allrfmo_month_5deg_longline_flag.rds")



data <- readRDS("rfmo_all_annual_gear_flag.rds") ### Data from JC's paper

#install.packages("sf")
library(sf)
lmpas <- st_read("clean_lmpas.gpkg") ### Data from JC's paper; this one is necessary for creating a spatial grid


## PROCESSING ##################################################################

# Find cells covered by MPAs ---------------------------------------------

#install.packages("terra")
library(terra)
# First, build some rasters
cells_1x1_within <-
  rast(nrows = 180, ncols = 360,
       xmin = -180, xmax = 180,
       ymin = -90, ymax = 90,
       nlyrs = 1, crs = "EPSG: 4326",
       resolution = 1, val = 1) %>%
  # Identify which cells are within an MPA polygon
  terra::extract(lmpas,
          xy = T,
          exact = T)  %>% # The `exact = T` argument calculates the percent cover
  mutate(grid = "1x1") %>%
  select(grid, lon = x, lat = y, fraction)

dim(cells_1x1_within) # There are 914 (781 without Johsnton) cells on the 1X1 grid that fall
# within an MPA. Note that not all these cells may have data reported by
# an RFMO and that not all are FULLY within an MPA... we'll get there

cells_5x5_within <-
  rast(nrows = 180, ncols = 360,
       xmin = -180, xmax = 180,
       ymin = -90, ymax = 90,
       nlyrs = 1, crs = "EPSG: 4326",
       resolution = 5, val = 1) %>%
  # Identify which cells are within an MPA polygon
  terra::extract(lmpas,
          xy = T,
          exact = T) %>%
  mutate(grid = "5x5") %>%
  select(grid, lon = x, lat = y, fraction)

dim(cells_5x5_within) # There are 97 (81 without Johnston) cells within an MPA using the 5x5 grid


cells_within <- bind_rows(cells_1x1_within,
                          cells_5x5_within) %>%
  filter(fraction >= 1) %>%
  select(-fraction)

# Build spatial grid -----------------------------------------------------------
# There are a few steps here:
# - First, build a grid of unique lat-lon combinations where RFMOs actually report
# some data
# - Then, remove cells that we've identified as being FULLY covered by an MPA
# - We then calculate the distance of each point to the nearest MP

#FLAG: This appears to be the first place where I'd add in my length data instead of the catch data JC uses

# Create a single feature to keep things tidy
lmpas_sfc <- lmpas %>%
  st_union()

#install.packages("units")
library(units)

# length dataframe does not have a grid column, but you could easily add that field and make it 1x1 for PS and 5x5 for LL
length_onlyPSLL <- length %>%
  filter(gear %in% c("longline", "purse_seine")) %>%
  mutate(grid = case_when(gear == "longline" ~ "5x5",
                          gear == "purse_seine" ~ "1x1",
                          TRUE ~ NA_character_))

length_onlyPSLL %>%
  filter(is.na(lon) | is.na(lat))

rfmo_spat_grid_length <- length_onlyPSLL %>%
  select(grid, lat, lon) %>%
  filter(!is.na(lon), !is.na(lat)) %>%
  distinct() %>%
  ### remove cells that fall fully within MPAs
  anti_join(cells_within, by = c("grid", "lat", "lon")) %>%
  ### convert to spatial object
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>% ### FLAG: try a more pacific-centered projection to see if it changes anything?
  #calculate distance to the nearest MPA
  mutate(dist = st_distance(., lmpas_sfc),
         dist = units::set_units(x = dist, value = "nautical_miles"), ### distances are now in nm
         dist = as.numeric(dist)) %>%
  ### keep only grid cells within 300 nautical miles of an MPA (note this is less than 600 that JC used but less clunky data is more gooder)
  filter(dist <= 300) %>%
  
  ##FLAG: stopped here 10-8 at 4pm; still need to wrap my head around this next bit of code for the treatments
  ### Build a series of dummies
  mutate(treatment_100 = assign_near_far(dist = dist, x = 100),
         treatment_200 = assign_near_far(dist = dist, x = 200),
         treatment_300 = assign_near_far(dist = dist, x = 300),
         near_100 = case_when(treatment_100 == "near" ~ 1,
                              treatment_100 == "far" ~ 0,
                              T ~ NA),
         near_200 = case_when(treatment_200 == "near" ~ 1,
                              treatment_200 == "far" ~ 0,
                              T ~ NA),
         near_300 = case_when(treatment_300 == "near" ~ 1,
                              treatment_300 == "far" ~ 0,
                              T ~ NA)) %>%
  # We now add information on which one is the nearest MPA
  st_join(lmpas, st_nearest_feature)

# Now convert it into a data frame
rfmo_grid_length <- rfmo_spat_grid_length %>%
  bind_cols(st_coordinates(.)) %>%
  st_drop_geometry() %>%
  rename(lon = X, lat = Y) %>%
  group_by(lat, lon) %>%
  mutate(id = cur_group_id()) %>%
  ungroup() %>%
  select(id, grid, lat, lon, wdpaid, name, dist, contains("treatment"), contains("near_"), ocean, year_enforced, month_enforced)




##````````` old 
# Make grid with distances and donuts ------------------------------------------
rfmo_spat_grid <- data %>%
  # Build distinct lat, lon pairs by grid type
  select(grid, lat, lon) %>%
  distinct() %>%
  # Remove cells identified as fully within an MPA
  anti_join(cells_within, by = c("grid", "lat", "lon")) %>% # This removes 110 grid cells that fall within an MPA and that have data
  # Convert into a spatial object
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  # Calculate distance to nearest MPA
  mutate(dist = st_distance(., lmpas_sfc),
         dist = units::set_units(x = dist, value = "nautical_miles"), # From now on, distances are in nautical mules
         dist = as.numeric(dist)) %>%
  # We'll keep only grid cells within 600 nautical miles of an MPA
  filter(dist <= 600) %>%
  # Build a series of dummies
  mutate(treatment_100 = assign_near_far(dist = dist, x = 100),
         treatment_200 = assign_near_far(dist = dist, x = 200),
         treatment_300 = assign_near_far(dist = dist, x = 300),
         near_100 = case_when(treatment_100 == "near" ~ 1,
                              treatment_100 == "far" ~ 0,
                              T ~ NA),
         near_200 = case_when(treatment_200 == "near" ~ 1,
                              treatment_200 == "far" ~ 0,
                              T ~ NA),
         near_300 = case_when(treatment_300 == "near" ~ 1,
                              treatment_300 == "far" ~ 0,
                              T ~ NA)) %>%
  # We now add information on which one is the nearest MPA
  st_join(lmpas, st_nearest_feature)

# Now convert it into a data frame
rfmo_grid <- rfmo_spat_grid %>%
  bind_cols(st_coordinates(.)) %>%
  st_drop_geometry() %>%
  rename(lon = X, lat = Y) %>%
  group_by(lat, lon) %>%
  mutate(id = cur_group_id()) %>%
  ungroup() %>%
  select(id, grid, lat, lon, wdpaid, name, dist, contains("treatment"), contains("near_"), ocean, year_enforced, month_enforced)


## EXPORT ######################################################################
# saveRDS(object = rfmo_grid,
#         file = here("data", "processed", "distance_grid.rds"))
```


###Make panel
```{r}
# rfmo_grid_length is your spatial file, equivalent to distance_grid.rds

# Load data --------------------------------------------------------------------
distance_grid <- rfmo_grid_length
rfmo_data <- length_onlyPSLL

## PROCESSING ##################################################################

# X ----------------------------------------------------------------------------
combined <- rfmo_data %>%
  left_join(distance_grid, by = c("grid", "lon", "lat")) %>%
  mutate(id = paste(lat, lon, sep = "_"),
         event = year - year_enforced,
         post = 1 * (event >= 0))

##check_mt(combined, cutoff = 0) ##FLAG: check out what this does; I run into a vmem error so if it's not necessary keep commented out

#check_effort_gear(combined)

#test(combined)

## EXPORT ######################################################################

# X ----------------------------------------------------------------------------
# saveRDS(object = combined,
#         file = here("data", "processed", "annual_panel.rds"))

colnames(combined) #variables to look at: gear; species; length_cm; name (MPA names); dist (distance to MPAs in nm); near_100; near_200;
#near_300; post
table(combined$name)


  #   Chagos Marine Protected Area                  Coral Sea - NPZ         Galápagos Marine Reserve Motu Motiro Hiva (Salas y Gómez) 
  #                            423                           706295                            36666                            68503 
  #         Nazca-Desventuradas MP              Niue Moana Mahu MPA                        Palau NMS                Papahānaumokuākea 
  #                          12298                           626401                          6305862                           289433 
  # Phoenix Islands Protected Area  Pitcairn Islands Marine Reserve                     PRI (Jarvis)                       PRI (Wake) 
  #                        2960470                           336846                          1113947                           157232 
  #                  Revillagigedo 
  #                            254 

plot_data <- combined %>%
  filter(
    species %in% c("SKJ", "YFT", "BET", "ALB"),
    post %in% c(0, 1),
    !is.na(length_cm)
  ) %>%
  mutate(
    period = factor(
      post,
      levels = c(0, 1),
      labels = c("Before protection", "After protection")
    )
  ) %>%
  group_by(species, period) %>%
  summarise(
    n = n(),
    mean_length = mean(length_cm),
    se = sd(length_cm) / sqrt(n),
    .groups = "drop"
  )

ggplot(plot_data, aes(x = period, y = mean_length, fill = period)) +
  geom_col(width = 0.7) +
  geom_errorbar(
    aes(ymin = mean_length - se, ymax = mean_length + se),
    width = 0.15,
    na.rm = TRUE
  ) +
  facet_wrap(~ species) +
  labs(
    x = NULL,
    y = "Mean length (cm)",
    title = "Mean fish length before and after protection"
  ) +
  theme_classic() +
  theme(legend.position = "none")

plot_data <- combined %>%
  filter(
    species %in% c("SKJ", "YFT", "BET", "ALB"),
    post %in% c(0, 1),
    !is.na(length_cm)
  ) %>%
  mutate(
    period = factor(
      post,
      levels = c(0, 1),
      labels = c("Before protection", "After protection")
    )
  )

ggplot(plot_data, aes(x = period, y = length_cm, fill = period)) +
  geom_boxplot() +
  facet_wrap(~ species) +
  labs(
    x = NULL,
    y = "Length (cm)",
  ) +
  theme_classic() +
  theme(legend.position = "none")






### PIPA Longline bigeye pre protections
PIPA_LL_bet_length_BP <- combined %>%
  filter(name == "Phoenix Islands Protected Area",
         post %in% c(0, 1),
         species == "BET", ### Keep bigeye tuna only
         gear == "longline")%>% ### Focus on longline
  select(species, length_cm, post)

library(ggridges)
### Plot
ggplot(data = PIPA_LL_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(post))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)


PIPA_LL_bet_length_line <- combined %>%
  filter(name == "Phoenix Islands Protected Area",
         species == "BET",
         gear == "longline",
         !is.na(year),
         !is.na(length_cm)) %>%
  group_by(year) %>%
  summarise(mean_length = mean(length_cm),
            n = n(),
            .groups = "drop")

ggplot(PIPA_LL_bet_length_line, aes(x = year, y = mean_length)) +
  geom_line(aes(group = 1), color = "grey50") +
  geom_point(aes(color = "black"), size = 3) +
  geom_smooth(
    aes(group = 1),
    method = "lm",
    se = FALSE,
    color = "black",
    linetype = "dashed"
  ) +
  scale_color_manual(
    values = c("0" = "steelblue", "1" = "firebrick"),
    labels = c("0" = "Before protection", "1" = "After protection")
  ) +
  labs(
    x = "Sampling year",
    y = "Average length (cm)",
    color = NULL,
    title = "Phoenix Islands bigeye longline average length"
  ) +
  theme_classic()
```
























###Trying to adapt JC's code to start plotting
```{r}
# Source custom funcions -------------------------------------------------------
source("00_set_up.R")

img <- function(pic) {here("data", "raw", "gear_fish_pics", paste0(pic, ".svg"))}

# Load data --------------------------------------------------------------------
most_relevant_panel <- readRDS("annual_relevant_mpa_gears_estimation_panel.rds")

## VISUALIZE ###################################################################
# BACI plots (Panels A-D) ------------------------------------------------------
# We first build BACI plots showing the mean CPUE for each of the before, after,
# control, impact groups. There is one main plot for each gear, and then a plot
# for a "good" and a "bad" example.

# Panel A - For all purse seine data
ps_data <- most_relevant_panel %>%
  filter(!wdpaid == "555512151", # Chagos is reported in different CPUE units
         gear == "purse_seine") %>%
  rename(cpue = cpue_tot)

#delta_cpue(ps_data)

ps_baci_plot <- baci_plot(data = ps_data) +
  guides(color = "none") +
  theme(legend.position = "inside",
        legend.position.inside = c(0, 0),
        legend.justification = c(0, 0)) +
  labs(y = "CPUE (MT / set)",
       fill = "Distance") +
  annotate(geom = "text",
           x = c(1.25, 1.95),
           y = c(31, 25),
           label = c("Change in CPUE near: 22.5%",
                     "Change in CPUE far: 8.7%"),
           color = c(gear_palette["PS"], "gray50"))
ps_baci_plot
# Panel B "spillover gradient" -------------------------------------------------
# Now we build plots where we show the change in CPUE as one moves away from the
# MPA border. We also build one for PS, one for LL, and then one for a good and
# a bad example.
ps_delta_cpue_dist_data <- most_relevant_panel %>%
  filter(!wdpaid == "555512151", # Chagos is reported in different CPUE units!wdpaid == "555512151",
         gear == "purse_seine",
         !is.na(near_100)) %>%
  group_by(id, wdpaid, name, dist, lat, lon, event, near_100, post) %>%
  summarize(effort = sum(effort),
            tot_mt = sum(tot_mt),
            .groups = "drop") %>%
  ungroup() %>%
  mutate(cpue = tot_mt / effort,
         dist = (floor(dist / 25) * 25) + 12.5,
         dist_f = as.factor(-1 * dist))

all_ps_delta_cpue_dist_plot <- ps_delta_cpue_dist_data %>%
  group_by(post, dist) %>%
  summarize(cpue = mean(cpue, na.rm = T), .groups = "drop") %>%
  ungroup() %>%
  pivot_wider(names_from = post,
              values_from = cpue, names_prefix = "cpue_") %>%
  mutate(delta = cpue_1 - cpue_0,
         pct_change = delta / cpue_0) %>%
  ggplot(aes(x = dist, y = delta)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 100, linetype = "dotted") +
  annotate(geom = "text", x = c(50, 150), y = 10, label = c("Near", "Far")) +
  geom_smooth(method = "loess", span = 0.9,
              fill = unname(gear_palette)[1],
              color = unname(gear_palette)[1]) +
  geom_point(shape = 21, size = 2,
             fill = "#08519B",
             color = "black") +
  scale_size_continuous(labels = scales::percent) +
  labs(x = "Distance from border (NM)",
       y = "Change in CPUE (MT / set)")

# Combine the panels -----------------------------------------------------------
p <- plot_grid(ps_baci_plot,
               all_ps_delta_cpue_dist_plot,
               labels = "auto")
p
```







### Regressions
```{r}
# Source custom functions ------------------------------------------------------
source("00_set_up.R")

# Load data --------------------------------------------------------------------
annual_panel <- readRDS("annual_full_estimation_panel.rds")
most_relevant_panel <- readRDS("annual_relevant_mpa_gears_estimation_panel.rds")

# SOME CHECKS BEFORE PROCEEDING TO ANALYSIS
# There should be 9 LSMPAS in the main regression panel for purse seine
length(unique(annual_panel[annual_panel$gear == "purse_seine",]$name)) == 9
# There should be 6 LMPSAS int he most relevant panel for purse seine
length(unique(most_relevant_panel[most_relevant_panel$gear == "purse_seine",]$name)) == 6

# ## PROCESSING ##################################################################
# Estimate DiD with all the data -----------------------------------------------
# Step 1: no fixed effects whatsoever.
install.packages("fixest")
library(fixest)
step1 <- feols(log(cpue_tot) ~ post + near + post:near,
               panel.id = ~id + year,
               data = annual_panel,
               subset = ~gear == "purse_seine",
               vcov = "IID")

# Step 2: Control for unobserved characteristics in space with grid-id fixed effects
step2 <- feols(log(cpue_tot) ~ post + near + post:near | id,
               panel.id = ~id + year,
               data = annual_panel,
               subset = ~gear == "purse_seine",
               vcov = conley(cutoff = 200))

# Step 3: Now add fixed effects by fleet to account for unobserved fleet-specific characteristics
step3 <- feols(log(cpue_tot) ~ post + near + post:near | id + flag,
               panel.id = ~id + year,
               data = annual_panel,
               subset = ~gear == "purse_seine",
               vcov = conley(cutoff = 200))

# Step4: Now add fixed effects effects by mpa-gear-year to account for time varying fishery-specific changes
step4 <- feols(log(cpue_tot) ~ post + near + post:near | id + flag + wdpaid ^ year,
               panel.id = ~id + year,
               data = annual_panel,
               subset = ~gear == "purse_seine",
               vcov = conley(cutoff = 200))

step5 <- feols(log(cpue_tot) ~ post + near + post:near + oni_avg:wdpaid + I(oni_avg ^ 2):wdpaid | id + flag + wdpaid,
               panel.id = ~id + year,
               data = annual_panel,
               subset = ~gear == "purse_seine",
               vcov = conley(cutoff = 200))

# Estimate DiD with subsamples of relevant MPA-gear combinations ---------------
# Main specification
# Step 1: no fixed effects whatsoever.
step1_r <- feols(log(cpue_tot) ~ post + near + post:near,
                 panel.id = ~id + year,
                 data = most_relevant_panel,
                 subset = ~gear == "purse_seine",
                 vcov = "iid")

# Step 2: Control for unobserved characteristics in space with grid-id fixed effects
step2_r <- feols(log(cpue_tot) ~ post + near + post:near | id,
                 panel.id = ~id + year,
                 data = most_relevant_panel,
                 subset = ~gear == "purse_seine",
                 vcov = conley(cutoff = 200))

# Step 3: Now add fixed effects by fleet to account for unobserved fleet-specific characteristics
step3_r <- feols(log(cpue_tot) ~ post + near + post:near | id + flag,
                 panel.id = ~id + year,
                 data = most_relevant_panel,
                 subset = ~gear == "purse_seine",
                 vcov = conley(cutoff = 200))

# Step4: Now add fixed effects effects by mpa-gear-year to account for time varying fishery-specific changes
step4_r <- feols(log(cpue_tot) ~ post + near + post:near | id + flag + wdpaid ^ year,
                 panel.id = ~id + year,
                 data = most_relevant_panel,
                 subset = ~gear == "purse_seine",
                 vcov = conley(cutoff = 200))

# Step 5: Add quadratic ENSO terms
step5_r <- feols(log(cpue_tot) ~ post + near + post:near + wdpaid:oni_avg + wdpaid:I(oni_avg ^ 2) | id + flag + wdpaid,
                 panel.id = ~id + year,
                 data = most_relevant_panel,
                 subset = ~gear == "purse_seine",
                 vcov = conley(cutoff = 200))

## Extract gear-specific coefficients ------------------------------------------
gear_stats <- step4_r %>%
  broom::tidy(.id = "sample", conf.int = TRUE, conf.level = 0.95) %>%
  filter(str_detect(term, ":")) %>%
  mutate(sample = "PS") %>%
  mutate(sample = str_remove(sample, ".+; sample: "),
         gear = str_sub(sample, 1, 2),
         mpa = str_remove(sample, "LL |PS "),
         mpa = fct_reorder(mpa, estimate, max),
         gear = ifelse(gear == "PS", "Purse seine", "Longline")) %>%
  filter(!sample == "Full sample")


## EXPORT ######################################################################

main_reg <- list(step1,
                 step2,
                 step3,
                 step4,
                 step5)

relevant_mpa_gear_reg <- list(step1_r,
                              step2_r,
                              step3_r,
                              step4_r,
                              step5_r)

# Export models ----------------------------------------------------------------
# Main models
# saveRDS(object = main_reg,
#         file = here("data", "output", "main_reg.rds"))
# 
# # Relevant MPA-gear combinations
# saveRDS(object = relevant_mpa_gear_reg,
#         file = here("data", "output", "relevant_mpa_gear_reg.rds"))
# 
# # Export table of MPA and gears ------------------------------------------------
# saveRDS(object = gear_stats,
#         file = here("data", "output", "relevant_mpa_gear_combination_model_coefs.rds"))
```

