### Below you will find the code associated with "Marine Protected Areas improve age structure of tropical Pacific tunas" by Jaskiel et al.
### All code and data can be found on GitHub (https://github.com/jaskielj/MPA-Effects-on-Tropical-Tuna-Size-Structure) and Zenodo ()

### Set Directory, Install/library packages, & bring in data
```{r}
setwd("/Users/Jacob/Documents/MPA Effects on Tuna Size Structure")
getwd()

library(tidyverse)
library(ggridges)

length <- read_rds("wcpfc_length_data.rds")
catch_purse_seine <- read_rds("allrfmo_month_1deg_purseseine.rds")
catch_longline <- read_rds("allrfmo_month_5deg_longline_flag.rds")
```

### Hawaii Catch - Longline
```{r}
### Bigeye
hawaii_bet <- length |>
  filter(between(lon, -165, -145), # Filter around Hawaii
         between(lat, 15, 25),
         year > 2000, # Keep 2000 > records
         species == "BET", # Keep bigeye tuna only
         gear == "longline") |> # Focus on longline
  mutate(yq = paste(year, quarter)) |>
 # uncount(freq) |>
  select(yq, species, length_cm)

### geartypes are handline_large, handline_small, longline, pole_line, purse_seine, ringnet, troll 

### Plot them
#png("BET_hawaii.png", units = "in", width = 6, height = 5, res = 1200)
ggplot(data = hawaii_bet,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()

# Hov-moller-style
hawaii_bet |>
  mutate(length_10 = floor(length_cm / 10) + 5) |>
  count(yq, length_10) |>
  group_by(yq) |>
  mutate(frac = n / sum(n)) |> # Normalize by quarter
  ggplot(mapping = aes(x = length_10, y = yq, fill = frac)) +
  geom_tile()


### Yellowfin
hawaii_yft <- length |>
  filter(between(lon, -165, -145), # Filter around Hawaii
         between(lat, 15, 25),
         year > 2000, # Keep 2000 > records
         species == "YFT", # Keep yft only
         gear == "longline") |> # Focus on longline
  mutate(yq = paste(year, quarter)) |>
 # uncount(freq) |>
  select(yq, species, length_cm)
 
### Plot them
#png("YFT_5deg_equator.png", units = "in", width = 6, height = 5, res = 1200)
ggplot(data = hawaii_yft,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()

# Hov-moller-style
hawaii_yft |>
  mutate(length_10 = floor(length_cm / 10) + 5) |>
  count(yq, length_10) |>
  group_by(yq) |>
  mutate(frac = n / sum(n)) |> # Normalize by quarter
  ggplot(mapping = aes(x = length_10, y = yq, fill = frac)) +
  geom_tile()


### Skipjack
hawaii_skj <- length |>
  filter(between(lon, -178, -155), # Filter around Hawaii
         between(lat, -15, 20),
         year > 2000, # Keep 2000 > records
         species == "SKJ", # Keep bigeye tuna only
         gear == "purse_seine") |> # Focus on longline
  mutate(yq = paste(year, quarter)) |>
 # uncount(freq) |>
  select(yq, species, length_cm)
       

### Plot them
#png("SKJ_purse_seine_S261-S287_Cruisetrack.png", units = "in", width = 6, height = 5, res = 1200)
ggplot(data = hawaii_skj,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
#dev.off()

# Hov-moller-style
hawaii_skj |>
  mutate(length_10 = floor(length_cm / 10) + 5) |>
  count(yq, length_10) |>
  group_by(yq) |>
  mutate(frac = n / sum(n)) |> # Normalize by quarter
  ggplot(mapping = aes(x = length_10, y = yq, fill = frac)) +
  geom_tile()
```




### Exploring Tuna Length in and around MPAs
```{r}
## Starting with PIPA (MPA only; see MPA + buffer zone below)

### PIPA Longline bigeye pre protections
PIPA_LL_bet_length_BP <- length |>
  filter(between(lon, -175.866, -169.706), ### Filter around PIPA
         between(lat, -6.618, -0.983),
         between(year, 2000, 2014), ### Keep records between 2000 and 2014
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### PIPA longline bigeye during protection
PIPA_LL_bet_length_DP <- length |>
  filter(between(lon, -175.866, -169.706), ### Filter around PIPA
         between(lat, -6.618, -0.983),
         between(year, 2015, 2022), ### Keep records between 2015 and 2022
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_bet_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### PIPA longline bigeye after protections lifted
PIPA_LL_bet_length_AP <- length |>
  filter(between(lon, -175.866, -169.706), ### Filter around PIPA
         between(lat, -6.618, -0.983),
         between(year, 2023, 2025), ### Keep records between 2015 and 2022
         species == "BET", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
```

### PIPA Bigeye from purse-seine
```{r}
### PIPA purse-seine bigeye pre protections
PIPA_PS_bet_length_BP <- length |>
  filter(between(lon, -175.866, -169.706), ### Filter around PIPA
         between(lat, -6.618, -0.983),
         between(year, 2000, 2014), ### Keep records between 2000 and 2014
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### PIPA purse-seine bigeye during protections
PIPA_PS_bet_length_DP <- length |>
  filter(between(lon, -175.866, -169.706), ### Filter around PIPA
         between(lat, -6.618, -0.983),
         between(year, 2015, 2022), ### Keep records between 2000 and 2014
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_bet_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### PIPA purse-seine bigeye after protections
PIPA_PS_bet_length_AP <- length |>
  filter(between(lon, -175.866, -169.706), ### Filter around PIPA
         between(lat, -6.618, -0.983),
         between(year, 2023, 2025), ### Keep records between 2000 and 2014
         species == "BET", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
```

### Yellowfin longline in PIPA
```{r}
### Yellowfin longline before protections
PIPA_LL_yft_length_BP <- length |>
  filter(between(lon, -175.866, -169.706), ### Filter around PIPA
         between(lat, -6.618, -0.983),
         between(year, 1953, 2014), ### Keep records between 2000 and 2014
         species == "YFT", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Yellowfin longline during protections
PIPA_LL_yft_length_DP <- length |>
  filter(between(lon, -175.866, -169.706), ### Filter around PIPA
         between(lat, -6.618, -0.983),
         between(year, 2015, 2022), ### Keep records between 2000 and 2014
         species == "YFT", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_yft_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Yellowfin longline after protections
PIPA_LL_yft_length_AP <- length |>
  filter(between(lon, -175.866, -169.706), ### Filter around PIPA
         between(lat, -6.618, -0.983),
         between(year, 2023, 2025), ### Keep records between 2000 and 2014
         species == "YFT", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
```

### Yellowfin purse-seine in PIPA
```{r}
### Yellowfin purse-seine before protections
PIPA_PS_yft_length_BP <- length |>
  filter(between(lon, -175.866, -169.706), ### Filter around PIPA
         between(lat, -6.618, -0.983),
         between(year, 1953, 2014), ### Keep records between 2000 and 2014
         species == "YFT", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Yellowfin purse-seine during protections
PIPA_PS_yft_length_DP <- length |>
  filter(between(lon, -175.866, -169.706), ### Filter around PIPA
         between(lat, -6.618, -0.983),
         between(year, 2015, 2022), ### Keep records between 2000 and 2014
         species == "YFT", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse_seine
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_yft_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Yellowfin purse-seine after protections
PIPA_PS_yft_length_AP <- length |>
  filter(between(lon, -175.866, -169.706), ### Filter around PIPA
         between(lat, -6.618, -0.983),
         between(year, 2023, 2025), ### Keep records between 2000 and 2014
         species == "YFT", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on purse-seine
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
```

### Skipjack longline (not major component but still looking) in PIPA 
```{r}
### Skipjack longline before protections
PIPA_LL_skj_length_BP <- length |>
  filter(between(lon, -175.866, -169.706), ### Filter around PIPA
         between(lat, -6.618, -0.983),
         between(year, 1953, 2014), ### Keep records between 2000 and 2014
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Skipjack longline during protections
PIPA_LL_skj_length_DP <- length |>
  filter(between(lon, -175.866, -169.706), ### Filter around PIPA
         between(lat, -6.618, -0.983),
         between(year, 2015, 2022), ### Keep records between 2000 and 2014
         species == "SKJ", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_skj_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Skipjack longline after protections
PIPA_LL_skj_length_AP <- length |>
  filter(between(lon, -175.866, -169.706), ### Filter around PIPA
         between(lat, -6.618, -0.983),
         between(year, 2023, 2025), ### Keep records between 2000 and 2014
         species == "SKJ", ### Keep bigeye tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_LL_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
```

### Skipjack purse-seine in PIPA
```{r}
### Skipjack purse-seine before protections
PIPA_PS_skj_length_BP <- length |>
  filter(between(lon, -175.866, -169.706), ### Filter around PIPA
         between(lat, -6.618, -0.983),
         between(year, 1953, 2014), ### Keep records between 2000 and 2014
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Skipjack purse-seine during protections
PIPA_PS_skj_length_DP <- length |>
  filter(between(lon, -175.866, -169.706), ### Filter around PIPA
         between(lat, -6.618, -0.983),
         between(year, 2015, 2022), ### Keep records between 2000 and 2014
         species == "SKJ", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_skj_length_DP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### Skipjack purse-seine after protections
PIPA_PS_skj_length_AP <- length |>
  filter(between(lon, -175.866, -169.706), ### Filter around PIPA
         between(lat, -6.618, -0.983),
         between(year, 2023, 2025), ### Keep records between 2000 and 2014
         species == "SKJ", ### Keep bigeye tuna only
         gear == "purse_seine") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PIPA_PS_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
```


## Palau National Marine Sanctuary

### PNMS Bigeye
```{r}
### Longline pre-protection
PNMS_LL_bet_length_BP <- length |>
  filter(between(lon, 129, 137), ### Filter around PIPA
         between(lat, 1, 12),
         between(year, 1953, 2019), ### Keep records between 2000 and 2014
         species == "BET", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_LL_bet_length_AP <- length |>
  filter(between(lon, 129, 137), ### Filter around PIPA
         between(lat, 1, 12),
         between(year, 2020, 2025), ### Keep records between 2000 and 2014
         species == "BET", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
```

### PNMS purse-seine Bigeye
```{r}
### Before protection
PNMS_PS_bet_length_BP <- length |>
  filter(between(lon, 129, 137), ### Filter around PIPA
         between(lat, 1, 12),
         between(year, 1953, 2019), ### Keep records between 2000 and 2014
         species == "BET", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_bet_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_PS_bet_length_AP <- length |>
  filter(between(lon, 129, 137), ### Filter around PIPA
         between(lat, 1, 12),
         between(year, 2020, 2025), ### Keep records between 2000 and 2014
         species == "BET", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_bet_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
```

### PNMS Yellowfin longline
```{r}
### Before protection
PNMS_LL_yft_length_BP <- length |>
  filter(between(lon, 129, 137), ### Filter around PIPA
         between(lat, 1, 12),
         between(year, 1953, 2019), ### Keep records between 2000 and 2014
         species == "YFT", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_LL_yft_length_AP <- length |>
  filter(between(lon, 129, 137), ### Filter around PIPA
         between(lat, 1, 12),
         between(year, 2020, 2025), ### Keep records between 2000 and 2014
         species == "YFT", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
```

### PNMS purse-seine yellowfin
```{r}
### Before protection
PNMS_PS_yft_length_BP <- length |>
  filter(between(lon, 129, 137), ### Filter around PIPA
         between(lat, 1, 12),
         between(year, 1953, 2019), ### Keep records between 2000 and 2014
         species == "YFT", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_yft_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_PS_yft_length_AP <- length |>
  filter(between(lon, 129, 137), ### Filter around PIPA
         between(lat, 1, 12),
         between(year, 2020, 2025), ### Keep records between 2000 and 2014
         species == "YFT", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_yft_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
```

### PNMS longline skipjack
```{r}
### Before protection
PNMS_LL_skj_length_BP <- length |>
  filter(between(lon, 129, 137), ### Filter around PIPA
         between(lat, 1, 12),
         between(year, 1953, 2019), ### Keep records between 2000 and 2014
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_LL_skj_length_AP <- length |>
  filter(between(lon, 129, 137), ### Filter around PIPA
         between(lat, 1, 12),
         between(year, 2020, 2025), ### Keep records between 2000 and 2014
         species == "SKJ", ### Keep skipjack tuna only
         gear == "longline") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_LL_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
```

### PNMS purse-seine skipjack
```{r}
### Before protection
PNMS_PS_skj_length_BP <- length |>
  filter(between(lon, 129, 137), ### Filter around PIPA
         between(lat, 1, 12),
         between(year, 1953, 2019), ### Keep records between 2000 and 2014
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_skj_length_BP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)

### After protection
PNMS_PS_skj_length_AP <- length |>
  filter(between(lon, 129, 137), ### Filter around PIPA
         between(lat, 1, 12),
         between(year, 2020, 2025), ### Keep records between 2000 and 2014
         species == "SKJ", ### Keep skipjack tuna only
         gear == "purse_seine") |> ### Focus on longline
  mutate(yq = paste(year, quarter)) |>
 ### uncount(freq) |>
  select(yq, species, length_cm)

### Plot
ggplot(data = PNMS_PS_skj_length_AP,
       mapping = aes(x = length_cm,
                     y = factor(yq))) +
  geom_density_ridges(vline_colour = "black",
                      vline_width = 1,
                      quantile_lines = T,
                      quantiles = 2)
```

